% !TeX root = ../../main.tex

\chapter{Concluding Remarks}
\label{chap:conclusion}

\section{Summary of Findings}

This dissertation explored the application of ultrasonic elastography techniques, specifically VisR and DoPIo, for noninvasive assessment of renal allograft health. The key findings from each chapter are as follows.

% Paragraph 1: VisR in humans

In Chapter~\ref{chap:visr}, VisR elastography was implemented in a cohort of renal transplant patients. The study demonstrated that VisR metrics exhibited modest correlations with IFTA severity as determined by histopathological analysis of renal biopsies. In terms of classifying various pathologies, VisR-derived DoA and RR distinguished mild and moderate IFTA with mean AUCs of around 0.69 despite limited sample sizes. These findings suggest that VisR has potential as a noninvasive tool for assessing renal allograft health, although major limitations related to study population, confounding factors, and assumptions inherent to the VisR technique were also showcased.

% Paragraph 2: DoPio development

Chapter~\ref{chap:dopio} introduced Double-Profile Intersection (DoPIo) elastography as an alternative analysis technique to extract more quantitative information from ARF-induced displacement profiles. The DoPIo method was validated in phantom studies, demonstrating its ability to quantify the shear elastic modulus of tissue-mimicking materials and \emph{ex vivo} tissues. Furthermore, DoPIo was able to quantify shear elasticity independently of ARF push magnitude, obviating a key complication associated with VisR.

% Paragraph 3: DoPIo vs VisR in pig IRI model

Finally, Chapter~\ref{chap:iri} compared the performance of VisR and DoPIo in a porcine ischemia-reperfusion injury (IRI) model. Both techniques were able to detect changes in renal parenchymal mechanics associated with IRI-induced inflammation and fibrosis. However, DoPIo provided more quantitative assessments of shear elasticity and outperformed VisR in detecting kidneys with the combined presence of inflammation and fibrosis, which may enhance its utility for specifically identifying graft pathologies like AMR that are relevant for monitoring graft health and survival.

% Integrated findings

Together, this dissertation demonstrated that both VisR and DoPIo elastography techniques can distinguish elastic changes in renal parenchyma associated with allograft dysfunction and rejection. As such, the results reported in this work support the hypothesis that the noninvasive assessment of microstructural changes in renal parenchyma using VisR and DoPIo can provide biologically relevant information about tissue microstructure that specifically correspond to inflammation and fibrosis.

\input{parts/conclusion/limitation}

\input{parts/conclusion/future}