% !TeX root = ../../main.tex

\chapter{INTRODUCTION}

\input{parts/intro/motivation}

\input{parts/intro/biomechanics}

\input{parts/intro/arfi}

\section{Hypothesis}

The underlying hypothesis of this dissertation is that VisR and DoPIo ultrasound detects changes in mechanical properties associated with inflammation and fibrosis in renal allografts \emph{in vivo}.

To evaluate this hypothesis, the objectives of this work are as follows:

1. Assess the feasibility of using VisR-derived qualitative metrics to noninvasively monitor human renal allografts \emph{in vivo}

2. Evaluate the ability to quantify tissue elasticity in stiffer materials and larger imaging depth ranges using DoPIo imaging

3. Demonstrate the potential to detect inflammation and/or fibrosis in porcine kidneys with \emph{in vivo} DoPIo-derived elasticity measures, and compare it to VisR-derived metrics

\section{Organization of This Dissertation}

This dissertation focuses on the implementation and evaluation of two ARF-based ultrasound elastography techniques, VisR and DoPIo, for noninvasive assessment of renal allograft health.

To this end, Chapter~\ref{chap:visr} presents the clinical implementation of VisR elastography in a cohort of renal transplant patients. VisR-derived metrics are evaluated for their ability to distinguish between healthy and dysfunctional grafts, as well as temporal changes in such metrics over time.

Chapter~\ref{chap:dopio} describes the development of DoPIo elastography, including the theoretical framework, simulation studies, and experimental validation in tissue-mimicking phantoms. The performance of DoPIo is characterized in terms of accuracy, precision, and optimal configurations.

Chapter~\ref{chap:iri} evaluates the ability of DoPIo-derived elasticity measures to detect inflammation and/or fibrosis in a porcine model of ischemia-reperfusion injury (IRI). VisR-derived metrics are also obtained for comparison. Histological analysis serves as the reference standard for assessing the presence and extent of inflammation and/or fibrosis.

Finally, Chapter~\ref{chap:conclusion} summarizes the key findings from the preceding chapters, discusses their implications for noninvasive monitoring of renal allograft health, and outlines potential directions for future research in this area.
