% !TeX root = ../main.tex

% Acknowledgements are the author's statement of gratitude to and
% recognition of the people and institutions who helped the author's
% research and writing.

\pretocheader{\bfseries ACKNOWLEDGEMENTS}

First, I would like to express my absolute, most sincere gratitude to my advisor, Prof. Caterina Gallippi. Her willingness to take a chance on me in my sophomore year of undergraduate studies -- as well as her continued provision of academic support, perspective regarding the ultrasound community, and encouragement towards creativity throughout my graduate career -- have been invaluable to my growth as a researcher and engineer. Her mentorship has been a cornerstone of my academic journey, especially during difficult personal tribulations, and I am eternally grateful for her guidance and patience.

I also wholeheartedly thank Asst. Prof. Md. Murad Hossain, now of the University of Hawai'i at M\={a}noa, for being a crucial mentor during my undergraduate years and the beginning of my graduate studies. His technical guidance and encouragement was instrumental in how I developed as a researcher, and I am thankful for the opportunities he provided me to grow my skills.

I would also like to offer my heartfelt appreciation to the members of my thesis committee. Prof. Gregg Trahey of Duke University has provided invaluable insights in developing my intuition around $k$-space and ultrasound imaging, both of which were fundamental to my transition from undergraduate to graduate research. Assoc. Prof. Gianmarco Pinton has been an incredible resource for my understanding of (non)linear systems and ultrasound simulation, and I am thankful for his willingness to share his expertise. Assoc. Prof. Emily Chang has provided essential support in my completion of the clinical aspects of this work, and I am grateful for her support and encouragement throughout my participation in trainee events at the UNC Kidney Center. Prof. Tim Nichols has also been a kind and reassuring face throughout the preclinical aspects of this dissertation, and I appreciate his willingness to provide feedback on my work. And lastly, Committee Chair and Asst. Prof. Kennita Johnson has been an incredible advocate for my success throughout my time at the Lampe Joint Department and a crucial leader-by-example for how academic engineers can still make practical and social impacts in their communities. For all of these committee members, I am deeply thankful for their time, expertise, and encouragement.

I also thank other key mentors, colleagues, and collaborators who have supported me throughout my academic journey. Dr. Melissa Caughey and Mrs. Melrose Fisher have been incredibly supportive figures who have added a distinctly personal and human touch to my graduate experience, and I am grateful for their encouragement, advice, and friendship. Prof. Randy Detwiler, Asst. Prof. Vanessa Moreno, Prof. Rani Sellers, and Dir. Gaby de la Cruz have provided critical feedback and histopathological insights that have strengthened the clinical relevance and soundness of this work, and I appreciate their willingness to share their expertise. Dr. Elizabeth Merricks and the rest of the Francis Owen Blood Research Laboratory team have been wonderful collaborators in the porcine study (and beyond), and I am thankful for their partnership and willingness to bring me into their fold. Of course, I would be remiss if I did not also thank my fellow Gallippi Lab members, past and present, for their camaraderie, technical insights, and moral support throughout my time here -- especially Dr.s Tomek Czernuszewicz and Chris Moore for helping me feel welcome from day one, Dr. Keerthi Anand for sticking with me through undergrad and as part of the same graduate cohort, Joey Richardson and Dr. Anna Phillips for being a source of many crucial inspirations and shifts in perspective, Sabiq Muhtadi for being an indispensable partner in the development of many parts of this work, and Matt Somerville for being an enthusiastic and considerate steward for what's to come.

I'd also like to acknowledge some additional key figures beyond my home institution and department. I am thankful for Prof.s Ruud van Sloun, Damien Garcia, Alfred Yu, Chih-Chung Huang, and Wei-Ning Lee, Yun Jing, Hong Chen, and Elisa Konofagou, as well as the ULTRASONC community, Caitlyn Smith, and Xufei Chen for helping me feel more welcome in the international ultrasound research community no matter the venue or company; Prof. Jeremy Dahl of Stanford University for allowing me to remotely audit his ultrasound physics course in 2021 (and therefore, technically fulfilling my high school dream of attending Stanford), Dr. Ron Daigle and Dr. Ole Marius Hoel Rindal for validating the salience of my technical skills and insights when I sorely needed it; Dr. Beka Layton, Dr. Leah Townsend, Prof. Jeff Warren, and Prof. Maryann Feldman for enabling career development opportunities around entrepreneurship and science/innovation policy that helped to solidify my career goals beyond academia; and Dr.s Blaide Woodburn, Sisi Tran, and Michaela Copp, Madigan Bedard, Alex Billingsley, and Emily Devericks, and Glory Dan-Dukor, as well as past and present members of the Graduate Business Careers Club at UNC-CH and its interinstitutional partners for arming me with essential opportunities in professional development that helped me plan my next steps forward. Whether the contributions lasted days or years, I'm appreciative of the outsized impact these individuals have had on my growth -- even if they may not have known it.

I'd like to express my utmost gratitude to my partner, family, and friends for their unwavering support throughout my academic journey. Their love, encouragement, understanding, and practical support have been my anchor through the challenges and triumphs of graduate school, and I am deeply thankful for their presence in my life.

And lastly, I'd like to thank you, the reader, for taking the time to engage with my work. If this dissertation or its derivative works provided you with any insights, inspiration, evidence, or some other practical value, then you've directly made my efforts worthwhile.
