%Acknowledgements are the author's statement of gratitude to and
%recognition of the people and institutions who helped the author's
%research and writing.

\begin{center}
\vspace*{52pt}
{ACKNOWLEDGMENTS}
\end{center}

This dissertation would not have been possible without the remarkable support I’ve received over the years at UNC. I am especially grateful to my advisor, Cameron Ballard-Rosa, whose unwavering encouragement and guidance continually pushed me to do my best work. Cameron was one of the main reasons I chose to attend UNC, and that decision proved invaluable. 

I also received support and mentorship from the other outstanding members of my committee: Chris Clark, Mark Crescenzi, Marc Hetherington, and Junghyun Lim. I'd like to thank Chris Clark for his consistently insightful feedback and for fostering a classroom environment where I could develop the ideas that shaped this dissertation. If all political scientists were as caring, positive, and thoughtful a mentor as Chris the field would be in a much better place. Marc Hetherington was immensely generous with both his time and his financial resources, sharing survey space for multiple components of my dissertation. He always challenged me to think deeper about the implications of my ideas. Junghyun Lim pushed me to think about the bigger picture and to never lose sight of the tremendous normative implications of climate research. Finally, I am deeply grateful for Mark Crescenzi, who has supported me in countless ways, both professionally and personally. Mark reminded me to keep perspective beyond academia and was always a source of encouragement and tremendous support.

Over the past several years, I have also benefited tremendously professionally and personally from the entire political science community at UNC. I would like to thank the participants at the International Relations Brown Bag Seminars. I am deeply grateful to all seminar participants, but particularly to Ashley Anderson, Navin Bapat, Stephen Gent, and Tyler Pratt who always had helpful comments and modeled excellent seminar behavior. I would also like to thank my friends in the department, who have helped push my work forward and supported me outside of graduate school: Ryan Brooks, Devin Case-Ruchala, Tyler Ditmore, Brian Forester, Bryce Hecht, and Victoria Staub. I am especially grateful to Tyler, as a friend, sounding-board, and coauthor of one of the papers in this dissertation; collaborating with him was an invaluable experience and helped me realize the benefit of truly collaborative work. Working with Tyler has greatly improved my own work and helped remind me that political science research can be fun. 

I would also like to thank my friends outside of academia who provided much-needed respite throughout this journey and patiently listened to far too many of my rants and raves. In particular, I am grateful to my friends at \textit{The Pub} discord server for always offering a welcoming space and a much-needed outlet. I also thank Luke Koch for sharpening my mind with our daily puzzle competitions, running our weekly Dungeons and Dragons sessions which was a much needed escape, and for his valuable copy edits. Furthermore, I am deeply appreciative of Lilly Richards and Leeor Hershkovich for their immense patience in sitting through countless practice presentations and for their constant support and encouragement.

Finally, I would like to thank my family, Elizabeth, Jim, and Stephanie, who were always supportive and generous with their time, reading and editing countless pages. They were also willing companions and conversation partners on endless walks, for which I am incredibly grateful. I also thank Pumpkin and Wicket for their endless love and support throughout graduate school. Finally, and most importantly, I owe an enormous debt to my fiancé, Amelia. Pursuing a Ph.D. alongside a partner doing the same was, at times, a challenging experience, but we consistently supported each other, pushing our work forward with compassion, patience, and love, ultimately strengthening our respective dissertations. Her unwavering support and countless acts of kindness and encouragement literally made this dissertation possible. I am endlessly grateful for her love, encouragement, and partnership in every sense of the word.



\clearpage
